---
title: "NFL prediction"
subtitle: "NFL analyst 2023 regular seasons"
author: 
  - Che-Yu Wang
thanks: "Code and data are available at: 
date: April 1st 2024
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(tidymodels)
library(modelsummary)
library(nflverse)
```


# Introduction





# Data 



```{r}
#| echo: false
#| message: false
#| warning: false
nfl_2023 <- read.csv(here::here("data/analysis_data.csv"))
nfl_2023_train <- nfl_2023 |> 
  filter(week<=9)
nfl_2023_test <- nfl_2023 |> 
  filter(week>9)


```

Talk more about it.



```{r}
#| label: tbl-modelresults
#| tbl-cap: "Model for the results"
#| echo: false
first_model_tidymodels <-
  linear_reg() |>
  set_engine(engine = "lm") |>
  fit(
    passing_epa ~ completions + attempts + passing_yards,
    data = nfl_2023_train
  )

second_model_tidymodels <-
  linear_reg() |>
  set_engine(engine = "lm") |>
  fit(
    passing_epa ~ completions + attempts + passing_yards + sacks,
    data = nfl_2023_train
  )
modelsummary(list("model_1" = first_model_tidymodels, "model 2" = second_model_tidymodels))

```


```{r}
#| echo: false
predicte_results <- second_model_tidymodels |>
  predict(new_data = nfl_2023_test) |>
  cbind(nfl_2023_test) |>
  select(recent_team, week, passing_epa, .pred) |>
  arrange(recent_team, week) |>
  rename(predicted_passing_epa = .pred)


  
```
'Talk way more about it



# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Ap

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0+ \beta_1 x_{1i} +\beta_2 x_{2i}
\end{align}




# Results
```{r}


ggplot(predicte_results, aes(x = passing_epa, y = predicted_passing_epa)) + 
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red")

```

# Discussion



\newpage



\newpage


# References


